Portable Multi- and Many-Core Performance for Finite-Difference/Element Codes; Application to the Free-Surface Component of NEMO (NEMOLite2D 1.0).

### 2017-09-14 docs/papers/nemolite2d_performance/referee_response.txt

We have added a few key lines of source code from the Momentum kernel
in section 3.3. We have also added text that explains that the
apparently poor performance of this kernel is down to the number of
division operations.  On the Ivy Bridge, division operations require
at least eight times as many cycles as a multiplication which can be
very significant.

The suggestion of a "Related Work" section is a good one and we have added
this just after the Introduction.

We attribute the better performance of the SSE version of the code to the
fact that the poor SIMD efficiency of the code is failing to amortize the
larger cost of the peel and remainder loops associated with the greater
vector width of AVX. We have expanded upon this in the text.

Carlos Osuna's specific comments:

(Top of page 8.) We have clarified why allocating larger arrays was
necessary in order to use the safe_address directive.

Changed "Intel does..." to "Intel can..." as requested.

Sec 2.1.5 - the grid is not attached to the field, rather each field
keeps a pointer back to the grid upon which it is defined. This approach
was chosen in order to support applications which use more than one grid.

Sec 2.1.8 -

### 2017-09-14 docs/papers/nemolite2d_performance/nemolite2d_perf.tex
nemolite2d-paper: clarify safe_address directive and change to ...Intel can...

Much of the optimisations we have performed have been informed by the
diagnostic output produced by either the Cray or Intel compilers. Many
of the NEMOLite2D kernels contain conditional statements. These
statements are there to check whether \textit{e.g.} the current grid
point is wet or neighbours a boundary point. A compiler is better able
to optimise such a loop if it can be sure that all array accesses
within the body of the loop are safe for every trip, irrespective of
the conditional statements. In its diagnostic output the Cray compiler
notes this with messages of the form:
\begin{verbatim}
A loop starting at line 448 would benefit 
from "!dir$ safe_address".
\end{verbatim}
In order that we could safely add such directives we altered the
GOcean infrastructure to allocate all field-data arrays with extents
greater than strictly required.  We were then able to safely add the
\texttt{safe\_address} before all of the loops where the Cray compiler
indicated it might be useful (the Momentum loops and some of the BC
loops).

@@

Although the Intel compiler does do in-lining when routines are in
separate source files, we have found (both here and in our previous
work~\citep{shallow_psykal}) that the extent of the optimisations it
performs is reduced if it first has to in-line a routine. For the
Intel-compiled Shallow code, manually inserting kernel code at the
site of the subroutine call increased performance by about 25\%.

>>>>

Many of the optimisations we have performed have been informed by the
diagnostic output produced by either the Cray or Intel compilers. Many
of the NEMOLite2D kernels contain conditional statements. These
statements are there to check whether \textit{e.g.} the current grid
point is wet or neighbours a boundary point. A compiler is better able
to optimise such a loop if it can be sure that all array accesses
within the body of the loop are safe for every trip, irrespective of
the conditional statements. In its diagnostic output the Cray compiler
notes this with messages of the form:
\begin{verbatim}
A loop starting at line 448 would benefit 
from "!dir$ safe_address".
\end{verbatim}
Originally, all fields were only allocated the bare minimum of storage
and the conditional statements within kernels prevented out-of-bounds
accesses, \textit{e.g.} at the edge of the simulation domain.  We
subsequently altered the GOcean infrastructure to allocate all
field-data arrays with extents greater than strictly required.  This
enabled us to safely add the \texttt{safe\_address} before all of
the loops where the Cray compiler indicated it might be useful (the
Momentum loops and some of the BC loops).

@@

Although the Intel compiler can perform in-lining when routines are in
separate source files, we have found (both here and in our previous
work~\citep{shallow_psykal}) that the extent of the optimisations it
performs is reduced if it first has to in-line a routine. For the
Intel-compiled Shallow code, manually inserting kernel code at the
site of the subroutine call increased performance by about 25\%.
